---
title: "Food deserts analysis"
author: "Maria"
format: 
  html:
    theme:
      light: cosmo
      dark: [cosmo, theme-dark.scss]
    code-fold: true
    code-summary: "Show the code"
    code-tools: true
---


```{r}
#| echo: false
#| output: false
#| include: false
library(tidyverse)
library(leaflet)
library(plotly)
library(gt)
library(reactable)
library(sf)

```

```{r}
#| echo: false
#| output: false
#| include: false
data_osm<-read_rds("/Users/e5028514/Desktop/Projects/food/data/osm_new/food_osm_all_postcode_sa_ses_ra.rds")

data_abr<-read_rds("/Users/e5028514/Desktop/Projects/food/data/temp/food_abr_all_postcode_ses_sa.rds")

#data_full<-read_rds("map_data_all_plus_vic_r.rds")
```

### Setup


| Food category | General category | Amenities category |
|---------------|------------------|--------------------|
| bakery        | department_store | bar                |
| beverages     | supermarket      | cafe               |
| butcher       | wholesale        | fast_food          |
| cheese        |                  | food_court         |
| chocolate     |                  | ice_cream          |
| coffee        |                  | pub                |
| confectionery |                  | restaurant         |
| convenience   |                  | kiosk mall         |
| deli          |                  |                    |
| dairy         |                  |                    |
| farm          |                  |                    |
| frozen_food   |                  |                    |
| greengrocer   |                  |                    |
| health_food   |                  |                    |
| ice_cream     |                  |                    |
| pasta         |                  |                    |
| pastry        |                  |                    |
| seafood       |                  |                    |
| spices        |                  |                    |
| tea           |                  |                    |
| water         |                  |                    |
| alcohol       |                  |                    |
| wine          |                  |                    |

: OSM features categories collected

Areas:

Victoria 
ACT 
New South Wales 
Northern Territory 
SouthAustralia 
Queensland 
Tasmania

### Number of food venues per postcode

Some locations cannot be allocated to postcode, but they do have geometry. These are usually locations at ferries or terminals

Such as these ones - overall 67 observations

```{r}
data_osm|>filter(is.na(postcode_2021))|>st_drop_geometry()|>
  select(
    osm_id, 
    name, 
    starts_with("addr:"),
    shop,amenity,
    sa2_code_2021, sa2_name_2021,
    sa4_code_2021, sa4_name_2021,
    ses,
    ra
    )|>
 reactable( defaultPageSize = 5)
  
```
 
 

```{r}
data_osm|>filter(!is.na(postcode_2021))|>st_drop_geometry()|>
  mutate(
    landuse=case_when(landuse=="retail"~1,
                      landuse=="commercial"~2,
                      landuse=="farmyard"~3,
                      landuse=="vineyard"~4,
                      landuse=="construction" ~5,
                      TRUE~0),
    shop=case_when(shop=="convenience"~1,
                      shop=="supermarket"~2,
                      shop=="alcohol"~3,
                      shop=="bakery"~4,
                      shop=="mall" ~5,
                      shop=="butcher"~6,
                      shop=="department_store"~7,
                      shop=="greengrocer"~8,
                      shop=="wine" ~9,
                      TRUE~0),
    
  )

```

```{r}
data_osm|>
  group_by(postcode_2021)|>
  summarise(
    number_venues=n(),
    
    #landuse
    land_retail=sum(landuse==1),
    land_commectial=sum(landuse==2)/2,
    land_farmyard=sum(landuse==3)/3,
    land_vineyard=sum(landuse==4)/4,
    land_construction=sum(landuse==5)/5,
    
    #shop type
        convenience=sum(landuse==1),
    supermarket=sum(landuse==2)/2,
    bakery=sum(landuse==4)/4,
    mall=sum(landuse==5)/5,
    butcher=sum(landuse==6)/6,
    department_store=sum(landuse==7)/7,
    greengrocer=sum(landuse==8)/8,
    wine=sum(landuse==9)/9,
    alcohol=sum(landuse==3)/3,
  )|>reactable(searchable=TRUE,
               resizable=TRUE,
               defaultPageSize = 20,
               showPageSizeOptions = TRUE,
               filterable = TRUE
               )
```


### SA2


```{r}
data_osm|>
  group_by(sa2_code_2021, sa2_name_2021)|>
  summarise(
    number_venues=n(),
    
    #landuse
    land_retail=sum(landuse==1),
    land_commectial=sum(landuse==2)/2,
    land_farmyard=sum(landuse==3)/3,
    land_vineyard=sum(landuse==4)/4,
    land_construction=sum(landuse==5)/5,
    
    #shop type
        convenience=sum(landuse==1),
    supermarket=sum(landuse==2)/2,
    bakery=sum(landuse==4)/4,
    mall=sum(landuse==5)/5,
    butcher=sum(landuse==6)/6,
    department_store=sum(landuse==7)/7,
    greengrocer=sum(landuse==8)/8,
    wine=sum(landuse==9)/9,
    alcohol=sum(landuse==3)/3,
  )|>reactable(searchable=TRUE,
               resizable=TRUE,
               defaultPageSize = 20,
               showPageSizeOptions = TRUE,
               filterable = TRUE
               )
```



### SA4


```{r}
data_osm|>
  group_by(sa4_code_2021, sa4_name_2021)|>
  summarise(
    number_venues=n(),
    
    #landuse
    land_retail=sum(landuse==1),
    land_commectial=sum(landuse==2)/2,
    land_farmyard=sum(landuse==3)/3,
    land_vineyard=sum(landuse==4)/4,
    land_construction=sum(landuse==5)/5,
    
    #shop type
        convenience=sum(landuse==1),
    supermarket=sum(landuse==2)/2,
    bakery=sum(landuse==4)/4,
    mall=sum(landuse==5)/5,
    butcher=sum(landuse==6)/6,
    department_store=sum(landuse==7)/7,
    greengrocer=sum(landuse==8)/8,
    wine=sum(landuse==9)/9,
    alcohol=sum(landuse==3)/3,
  )|>reactable(searchable=TRUE,
               resizable=TRUE,
               defaultPageSize = 20,
               showPageSizeOptions = TRUE,
               filterable = TRUE
               )
```

